---
title: "p8105_hw1_cy2522"
output: github_document
author: "Chu Yu"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1 

## Start with a data frame

```{r  Problem1 solutions}
# data frame
library(tidyverse)
set.seed(1)

df_1 = tibble(
num_1 = runif(10,0,5),
logic_1 = c(num_1 > 2),
char_1 = c("this", "is", "my", "first", "home", "work", "problem", "of", "data", "science" ),
fac_1 = factor(c("a", "b", "c", "d", "e", "f", "g", "h", "i", "x"))
)

# mean
mean(df_1$num_1)      
mean(df_1$logic_1)
mean(df_1$char_1)       
mean(df_1$fac_1)
# means of the character vector and the  factor vector are NA.The means of them are regarded as missing value because they can not be calculated.
summary(df_1)
```

## 'as.numeric' function

```{r as.numeric, eval = FALSE}
# as.numeric

as.numeric(df_1$num_1)
as.numeric(df_1$logic_1)
as.numeric(df_1$char_1)
as.numeric(df_1$fac_1)
```
  
## converting
```{r convert}
# convert the characters:

char_fac_1 <- as.factor(df_1$char_1)
char_fac_1  
## when converted into factors, the charactors does not change much. The order of words changed by the initial letter.  

fac_num_1 <- as.numeric(char_fac_1)
fac_num_1
## they change into numbers 

# convert the factors:

fac_char_1 <- as.character(df_1$fac_1)
fac_char_1
## the words looks the same, but change into charactors.

char_num_1 <- as.numeric(fac_char_1)
char_num_1
## the value can not be obtained, meaning that charactors can not be converted into numbers, but factors can.
```


# Problem 2

## start with a data frame
```{r data frame}
# set the data frame

set.seed(1)
x <- rnorm(1000)

set.seed(4)
y <- rnorm(1000)

logic_2 <- c(x + y > 0)
num_2 <- as.numeric(logic_2)
fac_2 <- as.factor(logic_2)

plot_df <- data.frame(x, y, logic_2, num_2, fac_2)



```

# description of the data frame

##  the size of the dataset 
`r dim(plot_df)`

##  mean and median
`r mean(plot_df$x)`
`r median(plot_df$x)`
## proportion
`r propTRUE_plot_df <- prop.table(table(plot_df$logic_2))`
`r propTRUE_plot_df`



## scatterplots
```{r}
library(ggplot2)

# scatterplot 1 -- logic vector
ggplot(plot_df, aes(x = x, y = y, colour = logic_2)) +
  geom_point() +
  scale_fill_manual(values = c("blue", "red"),
                    labels = expression("x + y < 0", "x + y > 0")) +
  labs( x = "x", y = "y", title = "scatterplot of plot_df filled with logic vector") +
  theme(legend.position = "top", legend.title = element_blank()) 

ggsave("scatterplots1_hw1.pdf", width = 7, height = 7)

# scatterplot 2 -- numeric vector
ggplot(plot_df, aes(x = x, y = y, colour = num_2)) +
  geom_point() +
  scale_fill_manual(values = c("blue", "red"),
                    labels = expression("x + y < 0", "x + y > 0")) +
  labs( x = "x", y = "y", title = "scatterplot of plot_df filled with numeric vector") +
  theme(legend.position = "top", legend.title = element_blank()) 

# scatterplot 3 -- factor vector
ggplot(plot_df, aes(x = x, y = y, colour = fac_2)) +
  geom_point() +
  scale_fill_manual(values = c("blue", "red"),
                    labels = expression("x + y < 0", "x + y > 0")) +
  labs( x = "x", y = "y", title = "scatterplot of plot_df filled with factor vector") +
  theme(legend.position = "top", legend.title = element_blank()) 

```

